SIR Model With Testing Flows and Isolation

### 2020-10-19 codes/ali_test_sir.Rmd
pix was added
adding the parametrization of s

@@

![](../inst/pix/sir_test1.jpg)

@@

```{r model inputs, message=FALSE}
params <- c(N0=1000000, beta=1,gamma=1/3, omega=0.25, t=0.01, 
             W_s=0.01, W_i=1, W_r=0.01, #weights
             P_s=0, P_i=0.5, P_r=0.5, #prob of waiting for being tested positive
            eta_w=0.02, eta_t=0.01 #isolation parameter   
            )
class(params) <- "params_pansim"
state_init <- c(S_u=params[["N0"]], S_n=0,
                I_u=1,I_n=0,I_p=0,I_t=0,
                R_u=0,R_n=0,R_p=0,R_t=0,
                N=0,P=0)
# Disease free equilibrium
Sn_dfe <-function(params){
  unpack(as.list(params))
  return((t*(1-P_s)*N0)/omega)
}
Su_dfe<- params[["N0"]]-Sn_dfe(params)

@@

# R0<1:
R0(params=params)
# R0>1:
R0(params=update(params,beta=2))
```

@@

### Contour plot of $R_0=1$ and interpretations 
```{r contour_R0, message=FALSE,warning=FALSE }
 # specify the ranges
 beta  <- 0.5  ## set beta high enough to allow R0>1 in worst case
 eta_w <- seq(0,1, length.out=5) 
 eta_t <- seq(0,1, length.out=5) 
 t     <- seq(0,0.01, length.out=10)
 omega <- seq(0.1,1 , length.out=10) #note omega must be > t

 df1 <- expand.grid(N0=params[["N0"]],beta=beta,gamma=params[["gamma"]],omega=omega,t=t,
                    W_s=params[["W_s"]],W_i=params[["W_i"]],W_r=params[["W_r"]],
                    P_s=params[["P_s"]],P_i=params[["P_i"]],P_r=params[["P_r"]],
                    eta_w=eta_w,eta_t=eta_t)

@@

which(df2[]<0)
```

@@

```{r show_contours, width=24,height=24,message=FALSE,warning=FALSE}
p1 <- ggplot(df2,aes(x=omega,y=t,z=R0))+ theme_bw() +
    theme(panel.spacing=grid::unit(0,"lines"))
(p1
    + geom_contour_filled(breaks=pretty(unique(df2$R0),5),
                          alpha=0.5) ## breaks=0.2)
    + geom_contour(breaks=1,alpha=0.5,colour="black")
    + facet_grid(eta_w~eta_t, labeller=label_special)
    + scale_x_continuous(expand=expansion(c(0,0)), n.breaks=3)
    + scale_y_continuous(expand=expansion(c(0,0)), n.breaks=3)
    + scale_fill_viridis_d(name=parse(text="R[0]"))
)
```

@@

Note that $\eta_t\leq \eta_w$, i.e., reported people have a lower transmission probability than awaiting individuals. The perfect isolation/quarantine occurs when $\eta_w=\eta_t=0$ (see the upper left corner panel). Thus, we only consider the lower-triangle panels in the plot [not sure how to extract them in ggplot?].
(1) Looking at the first column of panels, there seems to be a compensatory relation between $t$ and $\omega$. i.e., higher $t$ requires smaller $\omega$ to keep R0 lower than 1. Also, for a fixed $\omega$, increasing $\eta_w$ requires higher $t$ to achive the critical R0.

>>>>

## Objectives

1- What combination of testing intensity, $t$, and isolation, $\eta_w$ and $\eta_t$, will result in $R_0 <1$?

@@

![](../pix/sir_test1.jpg)

@@

```{r model inputs, message=FALSE}
params <- c(N0=1000000, beta=1,gamma=1/3, omega=0.25, t=0.01, 
             W_s=0.01, W_i=1, W_r=0.01, #weights
             P_s=0, P_i=0.5, P_r=0.5, #prob of waiting for being tested positive
            eta_w=0.02, eta_t=0.01, #isolation parameter 
            s=2 #eta_w/eta_t
            )
class(params) <- "params_pansim"
state_init <- c(S_u=params[["N0"]], S_n=0,
                I_u=1,I_n=0,I_p=0,I_t=0,
                R_u=0,R_n=0,R_p=0,R_t=0,
                N=0,P=0)
# Disease free equilibrium
Sn_dfe <-function(params){
  unpack(as.list(params))
  return((t*(1-P_s)*N0)/omega)
}
Su_dfe<- params[["N0"]]-Sn_dfe(params)

@@

# R0 expression with eta_w/eta_t factored
R02<-function(params){
    unpack(as.list(params))
    Sn_dfe<- Sn_dfe(params)
    Su_dfe<- N0-Sn_dfe
    F_i <-t*N0*W_i/(W_s*Su_dfe) #at DFE

    A2 <- gamma*(P_i*F_i*omega*Sn_dfe+(omega+gamma)*(gamma*Sn_dfe+F_i*N0))
    B2 <- P_i*F_i*omega*(gamma*Su_dfe+omega*N0)
    D <- gamma*(omega+gamma)*(gamma*Su_dfe+omega*N0)
    C2 <- beta/((N0*gamma*(gamma*(omega+gamma)+F_i*(gamma+omega*P_i)))*(omega+gamma))
    # s <- eta_w/eta_t

    # return((eta_w*A2+eta_t*B2+D)*C2)
    return((eta_t*(s*A2+B2)+D)*C2)
}

# R0<1:
R0(params=params)
# R0>1:
R0(params=update(params,beta=2))

# check
all.equal(R0(params=params),R02(params=params))
```

@@

### Contour plot of $R_0=1$ and interpretations approache 1
```{r contour_R0_1, message=FALSE,warning=FALSE }
 # specify the ranges
 beta  <- 0.5  ## set beta high enough to allow R0>1 in worst case
 eta_w <- seq(0,1, length.out=5)
 eta_t <- seq(0.001,0.5, length.out=5) 
 t     <- seq(0,0.01, length.out=10)
 omega <- seq(0.1,1 , length.out=10) #note omega must be > t
# s <- seq(1,2.5,length.out=10)

 df1 <- expand.grid(N0=params[["N0"]],beta=beta,gamma=params[["gamma"]],omega=omega,t=t,
                    W_s=params[["W_s"]],W_i=params[["W_i"]],W_r=params[["W_r"]],
                    P_s=params[["P_s"]],P_i=params[["P_i"]],P_r=params[["P_r"]],
                    eta_w=eta_w,eta_t=eta_t,s=params[["s"]])

@@

# which(df2[]<0)
```

@@

```{r show_contours_1, width=24,height=24,message=FALSE,warning=FALSE}
p1 <- ggplot(df2,aes(x=omega,y=t,z=R0))+ theme_bw() +
    theme(panel.spacing=grid::unit(0,"lines"))
(p1
    + geom_contour_filled(breaks=pretty(unique(df2$R0),5),
                          alpha=0.5) ## breaks=0.2)
    + geom_contour(breaks=1,alpha=0.5,colour="black")
    + facet_grid(eta_w~eta_t, labeller=label_special)
    + scale_x_continuous(expand=expansion(c(0,0)), n.breaks=3)
    + scale_y_continuous(expand=expansion(c(0,0)), n.breaks=3)
    + scale_fill_viridis_d(name=parse(text="R[0]"))
)
```

### Contour plot of $R_0=1$ and interpretations approache 2, parametrization $s=\eta_w/\eta_t$

```{r contour_R0_2, message=FALSE,warning=FALSE }
 # specify the ranges
 beta  <- 0.5  ## set beta high enough to allow R0>1 in worst case
 # eta_w <- seq(0,1, length.out=5) 
 eta_t <- seq(0.001,0.5, length.out=5) 
 t     <- seq(0,0.01, length.out=10)
 omega <- seq(0.1,1 , length.out=10) #note omega must be > t
s <- seq(1,2.5,length.out=10)

 df11 <- expand.grid(N0=params[["N0"]],beta=beta,gamma=params[["gamma"]],omega=omega,t=t,
                    W_s=params[["W_s"]],W_i=params[["W_i"]],W_r=params[["W_r"]],
                    P_s=params[["P_s"]],P_i=params[["P_i"]],P_r=params[["P_r"]],
                    eta_w=params[["eta_w"]],eta_t=eta_t,s=s)

df22<- data.frame(df11,R0=apply(df11,1,function(params_in)R02(params=params_in)),
                      Fi= apply(df1,1,function(params_in)F_i(state=state_dfe,params=params_in)))

```{r show_contours_2, width=24,height=24,message=FALSE,warning=FALSE}
p11 <- ggplot(df22,aes(x=omega,y=t,z=R0))+ theme_bw() +
    theme(panel.spacing=grid::unit(0,"lines"))
(p11
    + geom_contour_filled(breaks=pretty(unique(df22$R0),5),
                          alpha=0.5) ## breaks=0.2)
    + geom_contour(breaks=1,alpha=0.5,colour="black")
    + facet_grid(s~eta_t, labeller=label_special)
    + scale_x_continuous(expand=expansion(c(0,0)), n.breaks=3)
    + scale_y_continuous(expand=expansion(c(0,0)), n.breaks=3)
    + scale_fill_viridis_d(name=parse(text="R[0]"))
)
```

@@

Note that $\eta_t\leq \eta_w$, i.e., reported people have a lower transmission probability than awaiting individuals. The perfect isolation/quarantine occurs when $\eta_w=\eta_t=0$ (see the upper left corner panel). Thus, we only consider the lower-triangle panels in the plot [not sure how to extract them in ggplot?].
(1) Looking at the first column of panels, there seems to be a compensatory relation between $t$ and $\omega$. i.e., higher $t$ requires smaller $\omega$ to keep R0 lower than 1. Also, for a fixed $\omega$, increasing $\eta_w$ requires higher $t$ to achieve the critical R0.
(2) A sensible result that we are expecting is $\partial{R_0}/\partial{\omega}<0$.
